Algebraic File Synchronization: Adequacy and Completeness

### 2016-10-12 p2/def-commands.tex
New definition of commands

@@

\begin{mydef}[Regular commands]
Therefore we have three pieces of information related to each command,
and represent them using triples that contain
their input type ($[v_{in}]$), output value ($v_{out}$),
and the node to which they are applied ($n$):
\[ \caaaa{v_{in}}{v_{out}}{n}. \]
\end{mydef}

@@

\noindent
A command can only succeed if the original value at node $n$ has a type that matches
the input type of the command. If this is not the case, or if the resulting
filesystem no longer has the tree property, then we say that the command
{\em breaks} the filesystem, and assigns $\fsbroken$ to it.
Broken filesystems are considered equal, but not equal to any working filesystem.

So that we could reason about sequences of commands that break every filesystem, 
we introduce, in addition to the regular commands,
the command $\cbrk$ that simply breaks every filesystem.

@@

\begin{mydef}[Effect of commands]
We use a replacement operator of the form
$\FS_{[\valv/n]}$ to denote a filesystem derived from a non-broken $\FS$ 
by replacing its value at $n$ with $\valv$:
\[ \FS_{[\valv/n]}(m) =
   \begin{cases}
   \valv &\mbox{if~} m=n\\
   \FS(m) &\mbox{otherwise.}
   \end{cases}
\]
The effect of the commands is then as follows:
\begin{align*}
&\cbrk\aFS = \fsbroken \\
&\cxynv\aFS = 
   \begin{cases}
   \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} [\FS(n)]\neq[x]\\
   \fsbroken &\mbox{if~} \FS_{[y/n]} \mbox{~violates the tree property}\\
   \FS_{[y/n]} &\mbox{otherwise.}
   \end{cases}
\end{align*}
\end{mydef}

>>>>

% What is encoded in a command?

@@

% Break / empty command

\noindent
A command at times will return an error.
This happens if the encoded type of the original filesystem value does not match
what is actually in the filesystem, or if the modified filesystem no
longer has the tree property.
In these cases we say that the command {\em breaks} the filesystem,
and the function we use to model the command will be undefined on the filesystem.

So that we could reason about sequences of commands that break every filesystem, 
we also introduce the command $\cbrk$ which is not defined anywhere,
that is, which simply breaks every filesystem.

% Definition

\begin{mydef}[Filesystem commands]
In other words, we model filesystem commands as partial endofunctions on filesystems,
that is, functions that map filesystems to filesystems, but which
are not defined everywhere:
\[ \alpha: \setf \nrightarrow \setf \]
A command $\alpha$ is not defined on filesystems that it would break.
At times in these cases we write that the command returns the bottom element, $\fsbroken$.
We write $\cbrk$ for the empty partial function which is not defined anywhere.

We have three pieces of information defining a non-empty command,
which we represent using a triple. These are its
input type ($[v_\textrm{in}]$), output value ($v_\textrm{out}$),
and the node to which it is applied ($n$):
\[ \alpha = \caaaa{v_\textrm{in}}{v_\textrm{out}}{n}. \]
\end{mydef}

@@

@@

\begin{mydef}[Effect of commands]
We use a replacement operator of the form
$\FS_{[\valv/n]}$ to denote a filesystem derived from $\FS$ 
by replacing its value at $n$ with $\valv$:
\[ \FS_{[\valv/n]}(m) =
   \begin{cases}
   \valv &\mbox{if~} m=n\\
   \FS(m) &\mbox{otherwise.}
   \end{cases}
\]
The effect of the commands is then as follows:
\begin{align*}
&\cbrk\aFS = \fsbroken \\
&\cxynv\aFS = 
   \begin{cases}
   % \fsbroken &\mbox{if~} \FS=\fsbroken\\
   \fsbroken &\mbox{if~} [\FS(n)]\neq[x]\\
   \fsbroken &\mbox{if~} \FS_{[y/n]} \mbox{~violates the tree property}\\
   \FS_{[y/n]} &\mbox{otherwise.}
   \end{cases}
\end{align*}
\end{mydef}

### 2016-10-12 p2/def-filesystem.tex
New definition of commands

We model a working filesystem
using a function $\FS$ with a set of nodes (filesystem paths) $\setn$ as its domain,
and a set of possible contents or values $\setv$ as its codomain.
Alternatively, if the filesystem has encountered a problem and is no longer usable, 
we assign it the special value $\fsbroken$:
\begin{mydef}[Filesystem]
\begin{gather*}
\FS =
\begin{cases}
\setn \rightarrow \setv \\
\fsbroken
\end{cases}
\end{gather*}
\end{mydef}
In our model, $\setn$ 
serves as a namespace or ``skeleton'' for the filesystem.
It contains all possible nodes, including the ones 
where the file system contains no file or directory,
where in our model $\FS$ has a special value, $\empt\in\setv$.

There is an ancestor / descendant relation defined over the nodes,
which arranges them in a (disjoint) union of rooted directed trees,
and which can be derived from the following $\parent$ function.
Tao et al. \cite{TSR} describe a similar filesystem model, although
they also model inodes and restrict the filesystem to just a single tree.
\begin{mydef}[$\parent$]
The function $\parent(n)$ returns the parent node of $n$, or
returns $\topnode$ if $n$ is the root of a tree.
\end{mydef}

@@

\begin{mydef}[Equivalence by type: $\typeeq$]
We write $v_1\typeeq v_2$ for $v_1,v_2\in\setv$ if $v_1$ and $v_2$ are both files,
are both directories, or are both $\empt$.
In other words, if $\setf$ is the set of file values and $\setd$ is the set of directories, then
with $\{\empt\}$ they form a partition of $\setv$:
\[ \setv/{\typeeq} = \{\setd,\setf,\{\empt\}\}. \]
\end{mydef}

@@

The formal definition of the tree property is therefore
\begin{mydef}[Tree property]
\begin{align*}
\forall n\in\setn:{} &\FS(n) \neq \empt \\
&\quad\wedge \parent(n) \neq \topnode \\
&\Rightarrow \FS(\parent(n)) \in \setd. 
\end{align*}
\end{mydef}

>>>>

We model a working filesystem
using a function $\FS$ with a set of nodes (filesystem paths) $\setn$ as its domain,
and a set of possible contents or values $\setv$ as its codomain.
\begin{mydef}[Filesystems, $\setn$, $\setv$ and $\setf$]
\[ \FS: \setn \rightarrow \setv \]
In our model, $\setn$ 
serves as a namespace or ``skeleton'' for the filesystem.
It contains all possible nodes, including the ones 
where the file system contains no file or directory,
where in our model $\FS$ has a special value, $\empt\in\setv$.
We write $\setfs$ for the set of filesystems.
\end{mydef}

There is an ancestor / descendant relation defined over the nodes,
which arranges them in a (disjoint) union of rooted directed trees,
and which can be derived from the following $\parent$ function.
Tao et al. \cite{TSR} describe a similar filesystem model, although
they also model inodes and restrict the filesystem to just a single tree.
\begin{mydef}[$\parent$]
The partial function $\parent:\setn\nrightarrow\setn$
returns the parent node of $n$,
and is undefined if $n$ is the root of a tree.
\end{mydef}

@@

\begin{mydef}[Equivalence by type: $\typeeq$]
We write $v_1\typeeq v_2$ for $v_1,v_2\in\setv$ if $v_1$ and $v_2$ are both files,
are both directories, or are both $\empt$.
% TODO needed? v
In other words, if $\setf$ is the set of file values and $\setd$ is the set of directories, then
with $\{\empt\}$ they form a partition of $\setv$:
\[ \setv/{\typeeq} = \{\setd,\setf,\{\empt\}\}. \]
\end{mydef}

@@

The formal definition of the tree property is therefore
\begin{mydef}[Tree property]
\[ \forall n\in\setn: \FS(n) \neq \empt \Rightarrow \FS(\parent(n)) \in \setd \]
where $\parent(n)$ is defined.
\end{mydef}
